---
title: "lab6"
author: "Vienna Saccomanno"
date: "2/14/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Goals: 

- Read in spatial data (shapefiles)
- Simplify polygons (st_simplify)
- Add and transform projection (st_transform and st_crs)
- Create several finalized-ish maps in R
- Join spatial data (st_join)
- Find spatial intersections (st_intersect)
- Interactive plots with tmap
- Updating color schemes, base layers, etc.
- Plotting spatial lines
- Creating sf data from lat/lon coordinates

Awesome resource: 
*Geocomputation in R* by Robin Lovelace, available online: 
<https://geocompr.robinlovelace.net/>

```{r}
library(tidyverse)
library(sf)
library(tmap)
library(leaflet)
library(ggrepel)
library(ggspatial)
library(RColorBrewer)
library(raster)
```

###Useful information on file types (from gisgeography.com):

- **.shp** is a mandatory Esri file that gives features their geometry. Every shapefile has its own .shp file that represent spatial vector data. For example, it could be points, lines and polygons in a map.

- **.shx** are mandatory Esri and AutoCAD shape index position. This type of file is used to search forward and backwards.

- **.dbf** is a standard database file used to store attribute data and object IDs. A .dbf file is mandatory for shape files. You can open .DBF files in Microsoft Access or Excel.

- **.prj** is an optional file that contains the metadata associated with the shapefiles coordinate and projection system. If this file does not exist, you will get the error “unknown coordinate system”. If you want to fix this error, you have to use the “define projection” tool which generates .prj files.

- **.xml** file types contains the metadata associated with the shapefile. If you delete this file, you essentially delete your metadata. You can open and edit this optional file type (.xml) in any text editor.

- **.sbn** is an optional spatial index file that optimizes spatial queries. This file type is saved together with a .sbx file. These two files make up a shape index to speed up spatial queries.

- **.sbx** are similar to .sbn files in which they speed up loading times. It works with .sbn files to optimize spatial queries. We tested .sbn and .sbx extensions and found that there were faster load times when these files existed. It was 6 seconds faster (27.3 sec versus 33.3 sec) compared with/without .sbn and .sbx files.

- **.cpg** are optional plain text files that describes the encoding applied to create the shapefile. If your shapefile doesn’t have a cpg file, then it has the system default encoding. 

###Mapping Examples

####Example 1: Dammed California

Data: California Jurisdictional Dams

Accessed from: <https://hub.arcgis.com/datasets/98a09bec89c84681ae1701a2eb62f599_0/data?geometry=-150.074%2C31.096%2C-87.54%2C43.298&page=10>

"This dataset is a feature class identifying all dams currently under the jurisdiction of the Division of Safety of Dams (DSOD). The dataset is extracted from DSOD internal records and contains basic information about the dam including the type of construction, basic dimensions such as height, length, and maximum storage capacity; abbreviated owner information to identify the entity legally responsible for the dam; an assessment of the downstream hazard associated with the dam; an assessment of the current condition of the dam; and indication as to whether the dam is operating at a restricted storage level. Several dams span rivers that define county boundaries, so DSOD references the right abutment of the dam to identify the location of the structure and to associate it with a singular administrative subdivision of California."

Data: California eco-regions (EPA)

Accessed from: <https://www.epa.gov/eco-research/ecoregion-download-files-state-region-9>

a. Read in the California ecoregions data (layer "ca_eco"), select only the attribute for eco-region (US_L3NAME), rename that to "Region", simplify the polygons (for time) using st_simplify, and set the CRS:

```{r}
#read in data
ca_eco<- read_sf(".", layer="ca_eco" ) %>% #SF because getting simple features information. "ca_eco" calls all file types with this start!
  dplyr::select(US_L3NAME) %>%  #select just geometries + attributes
  rename(Region = US_L3NAME) %>% 
  st_simplify(dTolerance = 100) %>% #goes through a polygon and collpses points within "x" distance (m). Makes big datasets more managemble
  st_transform(crs=4326)#takes an existing projection and change it

#Check projections st_crs(ca_eco)
```


```{r}
ca_counties <-read_sf(".", layer = "california_county_shape_file")#no crs associated with this file - need one
  
st_crs(ca_counties)=4326
  
```

```{r}
ca_dams<-read_sf(".", layer= "California_Jurisdictional_Dams") %>% 
  rename(Condition = Condition_)

ca_dams$condition <- fct_relevel(ca_dams$Condition,
                                 "Fair","Satisfactory","Unsatisfactory","Poor")
```


```{r}
plot(ca_eco) #st_simplify make this go way faster
plot(ca_counties) #longer because base plot it to plot something different based on every attribute in the file.
```

Make a map with GGplot
```{r}
#splitting max number of colors in RColorBrewer. We have 13 ecoregions
color_count<-13

my_colors<- colorRampPalette(brewer.pal(10, "Set2"))(color_count) #first is how many colors to use and then how to split the selected ramp

#SF likes ggplt and tmap. Ca_eco is base layer. Can't tell that we simplified data about using "simplify"
ggplot(ca_eco)+
  geom_sf(aes(fill=Region),#how we create the map using spatial info
          color="NA", #get rid of boundaries
          show.legend=FALSE)+
  scale_fill_manual(values=my_colors) +
  geom_sf(data=ca_counties, #adding counties. need to tell exactly where to get data
          fill = "NA",
          color = "gray30",
          size=0.1)+
  geom_point(data=ca_dams, #add dams (dams has lat and long in datafram)
             aes(x=Longitude, y= Latitude),
             size=1,
             color= "gray10",
             alpha = 0.4)+
  theme_minimal()+
  coord_sf(datum = NA) #remove lat /log numbers



```

What about subsets and intersections of spatial information. Dams in the Sierra Nevada Ecoregion
```{r}
#Joing dam and eco region data. Filter eco region and join with dam data that is assoicated with the filtered geometry
sn <- ca_eco %>% 
  filter(Region == "Sierra Nevada") %>% #only pulling out geometry for Sierra nevada ecoregion
  st_join(ca_dams) #joins geometry with dams

# Then plot:
ggplot(sn) + #put main layer (sn) in first line bc calling throughout = cleaner code
  geom_sf(data = ca_counties, fill = "wheat3", color = "NA") + #first data are all county info with no boundaries
  geom_sf(fill = "lemonchiffon4", color = "NA") + #now adding SN but not in this line bc it is above. It goes there first
  geom_point(aes(x = Longitude, y = Latitude), size = 0.5, color = "red4") + #adding dam points in sn- still pulling from sn finlter from first line of code
  theme_void() +
  coord_sf(datum=NA) +
  labs(x = "", y = "", title = "CA Dams in Sierra Nevada Eco-Region")
```

###Example 3. Ecoregions for SB county
Can plot just pieces using st_intersection (for example, if we only want to plot eco-regions in Santa Barbara County), and crop graphing space with coord_sf() limits. 

```{r}
# Get just SB county
sb <- ca_counties %>% 
  filter(NAME == "Santa Barbara")
#View(sb) = 4 rows of data because of the islands. Each row is its own polygon.

# Clip eco-region spatial data to intersection with SB county:
eco_clip <- st_intersection(ca_eco, sb) #intersection between two polygons. One you call first (eco) is the one getting clipped to the second (sb)

# Plot that!
ggplot(eco_clip) + #eco_clip is base layer, will be called throughout
  geom_sf(data = ca_counties, fill = "gray90", color = "gray80", size = 0.2) + # First add gray California bc want to have other counties shown while highlighting SB county
  geom_sf(aes(fill = Region), color = "NA") + # ...then add eco-regions (clipped)
  scale_fill_manual(values = c("darkolivegreen2","darkolivegreen","gold2")) + # Change color scheme of discrete outcomes that exist is SB county
  coord_sf(xlim = c(-121,-119), ylim = c(33.5,35.5)) + # Crop plotting area - limit coordinates included in big map
  geom_point(aes(x = -119.6982, y = 34.4208), size = 2) + # Add a point for SB City to give context
  geom_text(x = -119.6982, y = 34.35, label = "Santa Barbara") + #add a title for SB
  theme_minimal() + # Update theme
  theme(legend.position = c(0.5,0.15)) +# Move the legend
  labs(x = "", y = "", title = "Santa Barbara County Eco-Regions")
```

####Example 4. Intro to interactive plots with tmap
- tmaps can do interactive and non-interactive plots
- power plotting tool if ya want interactivity

```{r}
# First, create a tmap object that we'll then make interactive. Syntax is the same as ggplot
map_sb_eco <- tm_shape(eco_clip) + #saying take this info and make it a tm object. Haven't told it what we want to plot yet.
  tm_fill("Region", palette = "RdPu", alpha = 0.5) + #fill ecocliips
  tm_shape(ca_counties)+
  tm_borders()
  #tm_polygons() # Use tm_polygons for fill + lines; but can just show fill or borders (tm_fill or tm_borders)!

# Check class
# class(map_sb_eco)

# View it (note: some bg layers can take a while...)
tmap_mode("view") #this makes it interactive viewing. Maintained when knitting!
map_sb_eco

#Can change base map leaflet::providers
#example of different basemap
tm_basemap("BasemapAT")+
  tm_shape(eco_clip)+
  tm_borders(col="white")

```

####Example 5. SB fault line data

Fault line data from California Dept. of Conservation:  <https://maps.conservation.ca.gov/geology/#datalist>

Separate fault line types syncline/anticline, certain/concealed, direction columns using tidyr::separate(). 

```{r}
fault_lines <- read_sf(".", layer = "GMC_str_arc") %>% 
  st_transform(crs = 4326) %>% #existing projection (there is a .prj file)
  separate(LTYPE, into = c("syn_ant", "certainty", "direction"), sep = ",") #breaking up column "LTYPE" in 3 new columns. Separator to break this all up is a comma. R tells you there are some scenarios when there weren't 3 things - R adds an NA all good. This will help us to visualize base on different outcomes

# Base plot:
plot(fault_lines) #plots all the different attributes

# All CA with faultlines ontop. 

ggplot() + #keeping top layer empty bc we will be changing it
  geom_sf(data = ca_counties, fill = "black", color = "NA") + 
  geom_sf(data = fault_lines, aes(color = syn_ant)) + #making color different based on the column syn_ant
  theme_dark()

# Limit to faults within SB polygon: 
sb_faults <- fault_lines %>% 
  st_intersection(sb) # this time we are piping through it. taking the fault lines data and finding where it intersects SB

# Plotting with ggplot and zooming in on SB
ggplot() +
  geom_sf(data = sb) +
  geom_sf(data = sb_faults, aes(color = syn_ant))

# Plotting with tmap to make it interactive
tmap_mode("view") #default once set to view
tm_basemap("CartoDB.DarkMatter") +
  tm_shape(sb) +
  tm_borders(col = "gray50", lwd = 2) +
  tm_shape(sb_faults) +
  tm_lines(col = "syn_ant", palette = c("orange","purple"), lwd = 2) #"lwd" = line width
```

####Example 6. Faceted maps - just how you'd expect it to work. In ggplot and with tm_facets.
- based on different levels of a variable
- generally faceted by year
- here looking at types of fault lines

```{r}

#Back to all of CA
ggplot() +
  geom_sf(data = ca_counties, fill = "black", color = "NA") +
  geom_sf(data = fault_lines, aes(color = syn_ant)) + #color dependant on what kind of fault line it is
  theme_dark() +
  facet_wrap(~syn_ant) # Choose variable to facet by - kind of fault line it is. Plotting CA 4 times because 4 facets

# Can also do this with tmap: 
tm_basemap("CartoDB.DarkMatter") +
  tm_shape(sb) +
  tm_borders(col = "gray50", lwd = 2) +
  tm_shape(sb_faults) +
  tm_lines(col = "syn_ant", palette = c("orange","purple"), lwd = 2) +
  tm_facets(by = "syn_ant") #this makes it interactive

```

####Example 7. Creating spatial data from latitude/longitude information - from scratch

-making spatial points - only makes points with points
-Make data sf objects


California Sensitive Shoreline Sites (CA DFW: <http://data-cdfw.opendata.arcgis.com/datasets/252b33ef5ce94e1d8fc4cad67731b277_0>)

"The purpose of the sensitive site layer is to provide knowledge to spill responders of the location of sensitive sites in order to protect them during a spill response."

Read in the data:
```{r}

ca_sites <- read_csv("cadfw_sensitive_sites.csv") #back to a normal csv, not spatial

```

```{r}
# Read in by longitude and latitude in CSV, and set CRS. Make R know this is spatial data

sites_sf <- st_as_sf(ca_sites, coords = c("LONDD","LATDD"), crs = 4326) #st_as_sf is the function and then tell R where the coordinates are so that it know where the spatial data are. Otherwise R has no idea these are spatial. Also no projection so we need to set it. Now has sticky geometry and can use sf!

# Then make a plot: 
ggplot() +
  geom_sf(data = ca_counties, fill = "gray40") +
  geom_sf(data = sites_sf, aes(color = PRI_CODE), size = 0.5) #because sticky, don't use geom_point. Color by priority code.

```

